---
title: "Homework 3"
author: "Team 1"
date: "October 29, 2020"
output: html_document
---

# 1. Data Exploration



# 2. Data Preparation



# 3. Build Models



# 4. Select Models



# Appendix


# Library

```{r}
# load required packages
library(ggplot2)
library(dplyr)
#library(tidyr)
library(corrplot)
library(MASS)
library(caret)
library(RCurl)
library(tidyverse)
library(pROC)
library(RCurl)
```

```{r import}
# Loading the data
git_dir <- 'https://raw.github.com/odonnell31/DATA621-HW3/master/data'
train_df = read.csv(paste(git_dir, "/crime-training-data_modified.csv", sep=""))
test_df = read.csv(paste(git_dir, "/crime-evaluation-data_modified.csv", sep = ""))
head(train_df)
```

# Data Exploration & Preparation


See a summary of each column in the train_df set
```{r train_dfing_data_summary}
# view a summary of all columns
summary(train_df)
```

```{r}
# Correlations 
cor_train = cor(train_df,  use = "na.or.complete")
#cor_train = cor(train_df[sapply(train_df, function(x) !is.factor(x))])
corrplot(cor_train)
```

```{r}
pairs(~ target + zn + indus
      + chas + nox + rm + age + dis + rad + tax + ptratio + lstat + medv, data = train_df)
```

```{r}
# look at the structure of the variables
str(train_df)
str(test_df)
```

The summary() function for the training and testing data sets indicates that there are no missing values in the data. 
The response variable "target" is binary with 1 indicates crime rate is above median cirme rate and 0 indicates crime rate is not above median crime rate. 

Let's observe how the target variable is effected by other factors:
1. The plot of "target" against "age" shows target equalling one (above median crime rate) increases as the proportion of owner-occupied units built prior to 1940 increaases; the boxplot further shows that a larger mean of proportions of owner-occupied units built prior to 1940 is assoicated with higher crime rate.
2. Plots of crime rate against pupil-teacher ratio indicate higher crime rate "1" is associated with higher pupil-teacher ratio.

```{r}
par(mfrow=c(2,2))
# plot response variable "target" against predictor variable "age" 
plot(train_df$age,train_df$target)
boxplot(age ~ target, train_df )
# plot response variable "target" against predictor variable "ptratio"
plot(train_df$ptratio,train_df$target)
boxplot(ptratio ~ target, train_df)
```

NA stuff
```{r}
has_NA = names(which(sapply(train_df, anyNA)))
has_NA
```
There are no NAs


# Modeling

## 1) Binary Logistic Regression

```{r}
# preliminary exploration glm models
glm(formula = target ~ age, family = binomial(), data = train_df)
glm(formula = target ~ ptratio , family = binomial(), data = train_df)
```

### All predictor model
```{r}
all_preds = glm(target ~ ., family = binomial, data = train_df)
summary(all_preds)
train_df$preds = ifelse(all_preds$fitted.values > 0.5, 1, 0)
# look at confusion matrix
cm = confusionMatrix(as_factor(train_df$preds), as_factor(train_df$target), positive = "1")
cm
```


```{r}
step_all_preds = stepAIC(all_preds)
summary(step_all_preds)
train_df$preds = ifelse(step_all_preds$fitted.values > 0.5, 1, 0)
# look at confusion matrix
cm = confusionMatrix(as_factor(train_df$preds), as_factor(train_df$target), positive = "1")
cm
```

### Try treating chas and rad as factors

```{r}
# Based on data dictionary in hw assignment pdf and looking at the df,
# chas and rad should probably be factors
train_df2 = cbind(train_df)
train_df2$chas = as.factor(train_df2$chas)
train_df2$rad = as.factor(train_df2$rad)
all_preds_fac = glm(target ~ ., family = binomial, data = train_df2)
summary(all_preds_fac)
train_df2$preds = ifelse(all_preds_fac$fitted.values > 0.5, 1, 0)
# look at confusion matrix
cm = confusionMatrix(as_factor(train_df2$preds), as_factor(train_df2$target), positive = "1")
cm
```

```{r}
step_all_preds_fac = stepAIC(all_preds_fac)
summary(step_all_preds_fac)
train_df2$preds = ifelse(step_all_preds_fac$fitted.values > 0.5, 1, 0)
train_df2$pred_proba = step_all_preds_fac$fitted.values
# look at confusion matrix
cm = confusionMatrix(as_factor(train_df2$preds), as_factor(train_df2$target), positive = "1")
cm
```

```{r}
hist(step_all_preds_fac$fitted.values, main= "Histogram of Predicted Probabilities", xlab="Predicted Probabilities")
```

```{r}
proc = roc(train_df2$target, train_df2$pred_proba)
plot(proc, asp=NA, legacy.axes=TRUE, print.auc=TRUE, xlab="Specificity")
```